An Embedding Framework for Consistent Polyhedral Surrogates

### 2022-01-11 poly-embed-journal/poly-embed-journal.tex
small edits in Section 4; added explicit proof of Thm 6 in Appendix C and small edits there.

To prove Theorem \ref{thm:link-main}, it now suffices to show that for any polyhedral $L$ embedding some $\ell$, there exists a \emph{separated} link $\psi$ with respect to $L$ and $\ell$.
This is given by Construction ~\ref{const:eps-thick-link} below.

\begin{theorem} \label{thm:thickened-separated}
  Let polyhedral surrogate $L:\reals^d \to \reals^\Y_+$ embed the discrete loss $\ell:\R\to\reals^\Y_+$.
  Then there exists $\epsilon_0 > 0$ such that, for all $0 < \epsilon \leq \epsilon_0$, Construction~\ref{const:eps-thick-link} yields an $\epsilon$-separated link with respect to $L$ and $\ell$.
\end{theorem}

@@

For \emph{(a)}, begin by linking each embedding point back to its original report.
Now we must determine $\psi(u)$ for non-embedding points.
%Symbolically, if $u^* \in \argmin_u \inprod{p}{L(u)} = \Gamma(p)$, then $\psi(u^*) \in \argmin_r \inprod{p}{\ell(r)} = \gamma(p)$.
The challenge is that we may have $u \in \Gamma(p) \cap \Gamma(p')$.
Because $u$ minimizes expected surrogate loss for both $p$ and $p'$, we must have $\psi(u) \in \gamma(p) \cap \gamma(p')$.
It is not \emph{a priori} clear that these sets even have a nonempty intersection.
We use elicitation results, discussed in \S~\ref{sec:min-rep-sets}, to show that for each $u \in \reals^d$ there exists $r \in \R$ such that $\Gamma_u \subseteq \gamma_r$, i.e. any $p$ satisfying $u \in \Gamma(p)$ also satisfies $r \in \gamma(p)$.
This implies that if $u \in \Gamma(p) \cap \Gamma(p')$, then $r \in \gamma(p) \cap \gamma(p')$, so we may safely choose $\psi(u) = r$.

For \emph{(b)}, we show that this link can be ``thickened'' by some positive $\epsilon$, as described next.
Consider an optimal report set, i.e. set of the form $U = \Gamma(p) = \argmin_u \inprod{p}{L(u)}$.
By indirect elicitation, $\psi$ is already correct on $U$.
Now, we ``thicken'' $U$ to obtain $U_{\epsilon} = \{u : \|u - U\| \leq \epsilon\}$.
Then we require that all points in $U_{\epsilon}$ are linked to some element of $\gamma(p) = \argmin_r \inprod{p}{\ell(r)}$.
For $\epsilon > 0$, this directly implies calibration.

@@

Given $u \in U \cap U' \cap \dots$, we use $\Psi(u)$ to denote the remaining legal choices for $\psi(u)$ after imposing the requirements for each such set $U,U'$, etc.
The key claim is that, for small enough $\epsilon > 0$, $\Psi(u)$ is nonempty: there exists at least one legal report to link to.
This claim (Lemma \ref{lemma:thick-empty}) follows because each $U$ is a closed convex set.
If all of their thickenings $U_{\epsilon}$ intersect, then so do the $U$ sets themselves, in which case indirect elicitation implies that there exists a legal choice.
Importantly, we show that a thickening of their intersection (using a larger $\epsilon$) contains the intersection of their thickenings.
It is also important that, by Lemma~\ref{lem:polyhedral-range-gamma}, there are only finitely many sets of the form $U = \Gamma(p)$.
This yields a single uniform smallest $\epsilon$ such that the key claim is true for all $u \in \reals^d$.
%Given a well-defined thickened link $\psi$, calibration is almost immediate because, given $p$, any report $u$ linking to a suboptimal $r = \psi(u)$ satisfies $\inf_{u^* \in \Gamma(p)} \|u^* - u\| \geq \epsilon$.

@@

The next lemma shows that if a subset of $\U$ intersect, then their corresponding report sets intersect as well.
\begin{lemma} \label{lemma:calibrated-pos}
  Let $\U' \subseteq \U$.
  If $\cap_{U\in\U'} U \neq \emptyset$ then $\cap_{U\in\U'} R_U \neq \emptyset$.
\end{lemma}
\begin{proof}
  Let $u \in \cap_{U\in\U'} U$.
  Then we claim there is some $r$ such that $\Gamma_u \subseteq \gamma_r$.
  This follows from Proposition \ref{prop:embed-trim}, which shows that $\trim(\Gamma) = \{ \gamma_r : r \in \R\}$.
  Each $\Gamma_u$ is either in $\trim(\Gamma)$ or is contained in some set in $\trim(\Gamma)$, by definition, proving the claim.

  For each $U \in \U'$, for any $p$ such that $U = \Gamma(p)$, we have in particular that $u$ is optimal for $p$, so $p \in \Gamma_u$, so $p \in \gamma_r$, so $r$ is optimal for $p$.
  This implies that $\phi(r)$, the embedding point, is optimal for $p$, so $\phi(r) \in U$.
  This holds for all $U \in \U'$, so $r \in \cap_{U\in\U'} R_U$, so it is nonempty.
\end{proof}
Lemma \ref{lemma:calibrated-pos} implies that there exists a $\psi$ such that $(L,\psi)$ indirectly elicits $\ell$: for every $u$ that is optimal for some $p$, let $\U' = \{U\in\U : u \in U\}$ be the optimal sets that contain it; choose $r$ from the nonempty set $\cap_{U \in\U'} R_U$; and set $\psi(u) = r$.

The main problem now is to prove a ``thickened'' analogue of Lemma \ref{lemma:calibrated-pos}.
Namely, Lemma \ref{lemma:thick-empty} will show that if $\epsilon$ is small enough, then the $\epsilon$-thickenings of all $U \in \U'$ intersect if and only if the $U$ sets themselves intersect.
Thus, if $u \in \cap_{U \in \U'} B(U,\epsilon)$, then $u \in \cap_{U \in \U'} U$, and Lemma \ref{lemma:calibrated-pos} gives some legal target report $\psi(u) = r \in \cap_{U \in \U'} R_U$.

@@

\begin{corollary} \label{cor:thick-intersect}
  There exists $\epsilon_0 > 0$ such that, for any $0 < \epsilon \leq \epsilon_0$, for any subset $\{U_j : j \in \mathcal{J}\}$ of $\U$, if $\cap_j U_j = \emptyset$, then $\cap_j B(U_j,\epsilon) = \emptyset$.
\end{corollary}
\begin{proof}
  For each subset, Lemma \ref{lemma:thick-empty} gives an $\epsilon_0 > 0$.
  We take the minimum over these finitely many choices.
\end{proof}

@@

Theorem \ref{thm:thickened-separated}, that Construction \ref{const:eps-thick-link} is a separated link function for small enough $\epsilon$, follows immediately from well-definedness (Theorem \ref{thm:small-eps-thick}) and the definitions.

>>>>

\bo{Say anything about proof of above theorem?}
To prove Theorem \ref{thm:link-main}, it now suffices to show that for any polyhedral $L$ embedding some $\ell$, there exists a \emph{separated} link $\psi$ with respect to $L$ and $\ell$.
This is given by Construction ~\ref{const:eps-thick-link} below.

\begin{restatable}{theorem}{thickenedseparated} \label{thm:thickened-separated}
  Let polyhedral surrogate $L:\reals^d \to \reals^\Y_+$ embed the discrete loss $\ell:\R\to\reals^\Y_+$.
  Then there exists $\epsilon_0 > 0$ such that, for all $0 < \epsilon \leq \epsilon_0$, Construction~\ref{const:eps-thick-link} yields an $\epsilon$-separated link with respect to $L$ and $\ell$.
\end{restatable}

@@

For \emph{(a)}, begin by linking each embedding point back to its original report.
Now we must determine $\psi(u)$ for non-embedding points.
%Symbolically, if $u^* \in \argmin_u \inprod{p}{L(u)} = \Gamma(p)$, then $\psi(u^*) \in \argmin_r \inprod{p}{\ell(r)} = \gamma(p)$.
The challenge is that we may have $u \in \Gamma(p) \cap \Gamma(p')$.
Because $u$ minimizes expected surrogate loss for both $p$ and $p'$, the link must satisfy $\psi(u) \in \gamma(p) \cap \gamma(p')$.
It is not even clear \emph{a priori} that these sets intersect.
We use the definition of embedding and elicitation results, discussed in \S~\ref{sec:min-rep-sets}, to show that for each such $u$ there exists $r \in \R$ such that $\Gamma_u \subseteq \gamma_r$, i.e. any $p$ satisfying $u \in \Gamma(p)$ also satisfies $r \in \gamma(p)$.
This implies that if $u \in \Gamma(p) \cap \Gamma(p')$, then $r \in \gamma(p) \cap \gamma(p')$, so we may safely choose $\psi(u) = r$.

For \emph{(b)}, we show that this link can be ``thickened'' by some positive $\epsilon$, as described next.
Consider an optimal report set, i.e. set of the form $U = \Gamma(p) = \argmin_u \inprod{p}{L(u)}$.
By indirect elicitation, $\psi$ is already correct on $U$.
Now, we ``thicken'' $U$ to obtain $U_{\epsilon} = \{u : \|u - U\| \leq \epsilon\}$.
Then we require that all points in $U_{\epsilon}$ are linked to some element of $\gamma(p) = \argmin_r \inprod{p}{\ell(r)}$.
For $\epsilon > 0$, this directly implies separation.

@@

Given $u \in U \cap U' \cap \dots$, we use $\Psi(u)$ to denote the remaining legal choices for $\psi(u)$ after imposing the requirements for each such set $U,U'$, etc.
The key claim is that, for small enough $\epsilon > 0$, $\Psi(u)$ is nonempty: there exists at least one legal report to link to.
This claim follows from a geometric result (Lemma \ref{lemma:thick-empty}) that, for all small enough $\epsilon$, a subset of thickenings $U_{\epsilon}$ intersect if and only if the $U$ sets themselves intersect.
When they do intersect, indirect elicitation implies that there exists a legal choice of link for the intersection of the thickenings.
It is also important that, by Lemma~\ref{lem:polyhedral-range-gamma}, for polyhedral surrogates there are only finitely many sets of the form $U = \Gamma(p)$.
This yields a single uniform smallest $\epsilon$ such that the key claim is true for all $u \in \reals^d$.
%Given a well-defined thickened link $\psi$, calibration is almost immediate because, given $p$, any report $u$ linking to a suboptimal $r = \psi(u)$ satisfies $\inf_{u^* \in \Gamma(p)} \|u^* - u\| \geq \epsilon$.

@@

The next lemma shows that if a subset of $\U$ intersect, then their corresponding report sets intersect as well.
\begin{lemma} \label{lemma:calibrated-pos}
  Let $\U' \subseteq \U$.
  If $\cap_{U\in\U'} U \neq \emptyset$ then $\cap_{U\in\U'} R_U \neq \emptyset$.
\end{lemma}
\begin{proof}
  Let $u \in \cap_{U\in\U'} U$.
  Our first claim is that there exists $r$ such that $\Gamma_u \subseteq \gamma_r$.
  This follows from Proposition \ref{prop:embed-trim}, which shows that $\trim(\Gamma) = \{ \gamma_r : r \in \R\}$.
  Each $\Gamma_u$ is either in $\trim(\Gamma)$ or is contained in some set in $\trim(\Gamma)$, by definition, proving the first claim.
  Our second claim is that $r \in \cap_{U\in\U'} R_U$, which proves the lemma.
  To prove the second claim, take any $U \in \U'$.
  There is some $p$ such that $U = \Gamma(p)$, and we have in particular $p \in \Gamma_u$.
  By the first claim, $p \in \gamma_r$.
  By definition of embedding, $p \in \gamma_r \implies \varphi(r) \in \Gamma(p) = U$, so $r \in R_U$.
\end{proof}
Lemma \ref{lemma:calibrated-pos} implies that there exists a $\psi$ such that $(L,\psi)$ indirectly elicits $\ell$: for each $u$, let $\U' = \{U\in\U : u \in U\}$ be the optimal sets that contain it; choose $r$ from the nonempty set $\cap_{U \in\U'} R_U$; and set $\psi(u) = r$.

The main problem now is to prove a ``thickened'' analogue of Lemma \ref{lemma:calibrated-pos} that extends this link to points $u$ that are up to $\epsilon$ far from an optimal set $U$.
Namely, Lemma \ref{lemma:thick-empty} will show that if $\epsilon$ is small enough, then the $\epsilon$-thickenings of all $U \in \U'$ intersect if and only if the $U$ sets themselves intersect.
Thus, if $u \in \cap_{U \in \U'} B(U,\epsilon)$, then $u \in \cap_{U \in \U'} U$, and Lemma \ref{lemma:calibrated-pos} gives some legal target report $\psi(u) = r \in \cap_{U \in \U'} R_U$.

@@

\begin{corollary} \label{cor:thick-intersect}
  There exists $\epsilon_0 > 0$ such that, for any $0 < \epsilon \leq \epsilon_0$, for any subset $\{U_j : j \in \mathcal{J}\}$ of $\U$, if $\cap_j U_j = \emptyset$, then $\cap_j B(U_j,\epsilon) = \emptyset$.
\end{corollary}
\begin{proof}
  For each subset, Lemma \ref{lemma:thick-empty} gives an $\epsilon_0 > 0$.
  We take the minimum over these finitely many subsets of $\U$.
\end{proof}

@@

Theorem \ref{thm:thickened-separated}, which we restate here, is now almost immediate.
\thickenedseparated*
%% Bo: copying the text in a comment here for reference
%\begin{restatable}{theorem}{thickenedseparated} \label{thm:thickened-separated}
%  Let polyhedral surrogate $L:\reals^d \to \reals^\Y_+$ embed the discrete loss $\ell:\R\to\reals^\Y_+$.
%  Then there exists $\epsilon_0 > 0$ such that, for all $0 < \epsilon \leq \epsilon_0$, Construction~\ref{const:eps-thick-link} yields an $\epsilon$-separated link with respect to $L$ and $\ell$.
%\end{restatable}
\begin{proof}
  We create $\psi$ using Construction \ref{const:eps-thick-link} with the $L_{\infty}$ norm.
  By Theorem \ref{thm:small-eps-thick}, in particular there exists an $\epsilon$ such that $\psi$ is well-defined everywhere.

  To prove separation, suppose $u$ and $p$ are give such that $d_{\infty}(u,U) \leq \epsilon$, where $U = \Gamma(p)$.
  Then in Construction \ref{const:eps-thick-link}, $\psi(u) \in \Psi(u) \subseteq R_U = \{r : \varphi(r) \in U\}$.
  By definition of embedding, $\varphi(r) \in U = \Gamma(p) \implies r \in \gamma(p)$.
  So we obtain $\psi(u) \in \gamma(p)$ whenever $d_{\infty}(u,\Gamma(p)) \leq \epsilon$, which proves $\epsilon$-separation of the link $\psi$.
\end{proof}
